---
title: "Covid19 vaccine study Final Spring 2024"
author: "ediaz16"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid19) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: No difference in the effectiveness of the Covid 19 vaccine in the various gender subgroups.

$H_A$: There is a difference in the effectiveness of the Covid 19 vaccine in the various gender subgroups.

## Method 
The Cat~Cat helper file is used for the various gender subgroups:  
Males/categorical  
Females/categorical  
LGBTQ/categorical  
Druggies/categorical  




## Females
##Graphical Descriptive Results

```{r}
dataF=filter(FinalData,sex=="F")
barchartGC(~infected+treatment,data=dataF)
```

```{r}
barchartGC(~infected+treatment,data=dataF,type="percent")
```


The barchart shows the females with Covid 19 and recieved the vaccine had less symtoms than the females with the placebo. The females that did not have covid19 shows no difference in those that had been vaccinated and not vaccinated.

##Numerical Results

```{r}
table1<-xtabs(~infected+treatment,data=dataF)
rowPerc(table1)
```

```{r}
colPerc(table1)
```


This table shows the vaccine has a lower infected rate than the placebo. For the females with covid 19 and vaccinated shows a better treatment outcome.For the females unvaccinated it shows a 60% with symptoms and 40% of females had symptoms that had been vaccinated. 


##Inferential Results

```{r}
chisq.test(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test shows the p- value is 0.002162. 

# Overall Results and Conclusions

